\textbf{ Spectrophotometric parallaxes with linear models:\\ Accurate distances for luminous red-giant stars }

### 2019-06-02 paper/response_to_referee.txt

This point is important. The Gaia parallax uncertainties are based on
a spacecraft model that is independent of the measurements themselves.
Therefore, cuts on parallax uncertainties are not cuts on parallax and
thus are not subject to the strong censure that we give to such cuts!
But the referee is correct that we have been very unclear about this
and we have attempted to fix this in the relevant paragraph towards the
end of Section 2. This has improved the paper.

The referee is completely correct that the detailed distribution of
stars in parallax, SNR, and log g is not identical in the training set
and the parent sample.  We don't have any reason to believe that this
biases our results. However, we do now show these distribution
differences in a newly added figure, Figure XXX. Certainly these
differences mean that we probably have uncertainties that, in detail,
vary with log g. HOGG DO THIS

    4 - Finally, it is important to provide the reader with some
    understanding of how in practice the data are driving the resulting
    parallaxes. This may be achieved by discussing the behaviour of the
    linear coefficients theta. How do their values, as estimated by
    expr. (3), vary from training set A and to training set B? What
    spectral ranges are most relevant and survive the L1 regularization
    process? In other words, the authors could provide a discussion
    similar to the one around Figure 5 of Ness et al did on the Cannon,
    but this time comparing sets A and B across the spectrum or over the
    smaller photometric subspace.

Absolutely! Actually we had a section about this but dropped it just
before submission. One of the reasons we dropped it is because the
story is somewhat complicated. But you are right that this must be
there, and we have added it back in, in the middle of the Discussion
section. The short story (which you can read in the updated text) is
that the photometric part of the model is interpretable and stable,
but the spectroscopic part is less stable in its internals (while
remaining very stable in its predictions in the data space). We don't
fully understand this, but it connects to known issues in machine
learning more generally, which we point out in the text.

@@

    Section 2, 1st par. after the list of assumptions. The statement "We
    never cut on parallax or parallax signal-to-noise." is repeated.

Yes, yes it is. Fixed and removed.

    Section 4: "The hope of the model assumptions is that the Training Set
    will contain sufficient dust variation that the model will naturally
    learn the dust corrections and extrapolate acceptably.” Is this hope
    checked in any way?

Hahaha no. Okay fine. We now comment on this point in the Discussion,
where we now deliver an interpretation of the photometric part of the
model, and comment on the fact that it *does* look dust-corrected.

### 2019-06-02 paper/sp.tex
two more items

\section*{~}\clearpage
\section{Introduction} \label{sec:intro}

@@

\section{Method}

@@

\section{Data}

@@

\section{Results and validation}

@@

\section{Discussion}

@@

Another  advantage of a linear model over more general methods (like deep learning, say),
is that it is possible to look inside
the linear model and check whether the dependencies represented there make sense.
For example, the feature vector $x_n$ for star $n$
contains a set of photometry in magnitudes.
Linear combinations of these photometric measurements are like complex synthetic
magnitudes or colors.
Similarly, linear combinations of spectral pixels are like a projection of the spectral space.
Brief inspection of these linear combinations showed them to be sensible, but it is out of
scope here to interpret them in detail.

>>>>

\section*{~}\clearpage
\section{Introduction}\label{sec:intro}

@@

\section{Method}\label{sec:method}

@@

\section{Data}\label{sec:data}

@@

\section{Results and validation}\label{sec:results}

@@

\section{Discussion}\label{sec:discussion}

@@

Another  advantage of a linear model over more general methods (like deep learning, say),
is that it is possible to look inside
the linear model and check whether the dependencies represented there make sense.
For example, the feature vector $x_n$ for star $n$
contains a set of photometry in magnitudes.
Linear combinations of these photometric measurements are like complex synthetic
magnitudes or colors.
Similarly, linear combinations of spectral pixels are like a projection of the spectral space.

For an example of how the model might be interpreted, we can look at
the photometric part of the linear model and see if the combination of
photometry does indeed look like a dust-corrected absolute magnitude,
as we expected (see \sectionname~\ref{sec:data}).
It does! Indeed the model automatically finds a photometric
combination that includes a term close to $H-W2$, which is known as a
good extinction estimator.
And we find that on the two-fold A/B split, the photometric model is extremely stable.
As for the spectroscopic-pixel part of the model, we find that interpretation is
less simple, for two reasons:
One is that the number of pixels is enormous, so the model has huge flexibility.
Another is that although the two splits, A and B, both give similar-performing
models, they have (in detail) different linear coefficients in the spectral space.
That is, the model is not as stable in the spectroscopic domain as it is in the
photometric domain.
This is a generic feature of high-capacity machine-learning methods.
And it means that we did not obtrain as much insight from the spectroscopic model
as we had hoped up-front.
